\documentclass[cronos]{ling-paper}
\usepackage{float}
\usepackage{caption}
\captionsetup{
  font=footnotesize,
  singlelinecheck=false,
  justification = centering,
  labelformat = parens,
  position = top,
  belowskip = 10pt
}
\addbibresource[location=remote]{/home/patrl/repos/bibliography/elliott_mybib.bib}

\title{Continuation semantics and the split-scope signature}
\subtitle{A note on Hirsch's (2017) argument that type-flexibility is insufficient}

\input{acronyms}
\renewcommand*{\acsfont}[1]{\textsc{#1}}

\author{Patrick D.\,Elliott\\Massachusetts Institute of Technology}

\usepackage{appendix,float}

\begin{document}

\maketitle

\citet{hirschThesis} argues that \textit{split scope} readings of conjunction
can't be properly accounted for by positing a type-flexible meaning for
\textit{and}, and conversely that \ac{cr} \textit{must} be available in order to
account for the relevant data. In this note, we show that this argument doesn't
go through -- concretely, \citeauthor{barker2002}'s analysis of conjunction
within \textit{continuation semantics} (see also \citealt[chapter
7]{barkerShan2015}) \textit{predicts} the split-scope signature, by virtue of
independently motivated type-shifting operations.

\section{Hirsch's argument}

Hirsch's argument revolves around the sentence in (\ref{hirsch1}). The key
observation is that (\ref{hirsch1}) has a reading which entails that there are
two things that John refused to do: (i) \textit{visit any city in Europe} and
(ii) \textit{visit any city in Asia}. Under this reading, \textit{and} scopes
above \textit{refuse}, but the material in each conjunct must scope below
\textit{refuse}, since each conjunct contains an \ac{npi} licensed by \textit{refuse}.

\ex \textsc{The split-scope signature}\\
John refused to visit any city in Europe and any city in
Asia.\label{hirsch1}\hfill(\citealt[p.\,90]{hirschThesis})\\
\phantom{,}\hfill $\wedge > \ml{refuse} > \ml{any}$
\xe

\citeauthor{hirschThesis} characterizes the \enquote{split-scope signature} as
follows:\footnote{As noted by \citeauthor{hirschThesis}, \citet{parteeRooth}
  already recognized the possibility split-scope readings of \textit{dis}junction.}

\ex
\textit{And} scopes above some operator, which the apparent \ac{dp} conjuncts
scope below.
\xe

\citeauthor{hirschThesis}'s account of the split-scope signature is to derive
(\ref{hirsch1}) from the \ac{lf} in (\ref{hirsch-lf}), via deletion mechanisms.
The details are unimportant for our purposes, but it should be clear enough how
the \ac{lf} in (\ref{hirsch-lf}), if available, accounts for the split-scope signature:

\ex
John $λ x$ $\begin{aligned}[t]
  &\text{[\sub{\&P} [\sub{$v$P$_{1}$} $x$ refused [ any city in Europe
    $\lambda y$ \textsc{pro} to visit $y$]}\\
  &\text{and [\sub{$v$P$_{2}$} $x$ refused [ any city in Asia $λ z$ \textsc{pro}
    to visit $z$]]]]}
  \end{aligned}$\label{hirsch-lf}
\xe

\citeauthor{hirschThesis} shows in detail that the following denotation for
\textit{and} (a special case of \citeauthor{parteeRooth}'s
\citeyear{parteeRooth} recursive definition) can't capture this reading, on the
assumption that (\ref{hirsch1}) in fact involves \ac{qp} conjunction:

\ex
$\eval{and\sub{Q}} ≔ λQ₁Q₂P . Q₁ P ∧ Q₂ P$\hfill$\type{((e → t) → t) → ((e → t) → t) → (e → t) → t}$
\xe

\section{Enter \textit{continuations}}

\subsection{Background}

\citeauthor{hirschThesis}'s argument takes the standard-bearer for a
type-flexible theory of conjunction to be \citeauthor{parteeRooth}'s recursive
definition. Subsequently, however, a generalization of
\citeauthor{parteeRooth}'s account has been proposed by \citet{barker2002} and
\citet{barkerShan2015} within the framework of \textit{continuation semantics}.
Continuation semantics provides a perspective on scope-taking that generalizes
\citeauthor{montague1973}'s account of quantification, taking inspiration from
work in computer science on using delimited continuations to model control flow
(see, e.g., \citealt{danvyFilinski1992,wadler1994}). For our purposes,
continuation semantics amounts to the following conjecture: composition of
\textit{scopal} meanings is accomplished \textit{in-situ}, mediated by three
type-flexible operations -- \textit{lift} ($↑$), \textit{lower} ($↓$), and
\ac{cfa} ($\ml{S}$).

The definitions of lift/lower are straightforward, so we provide those first.
Lift, defined in (\ref{def:lift}), is simply a type-flexible version of \citeauthor{partee1986}'s
(\citeyear{partee1986}) Montague Lift. Lower, defined in (\ref{def:lower}) is
simply an instruction to feed a function $m$ the identity function over
truth-values as its argument. In a continuized fragment, the role of lift is to
allow non-scope-takers to compose with scope-takers, and the role of lower is to
get an ordinary value back from a scopal value.

\pex
\a Lift (def)\\
$a^{↑}  ≔ λ k . k a$\hfill$↑:\type{a → (a → t) → t}$\label{def:lift}
\a Lower (def)\\
$m^{↓} ≔ m (λx . x)$\hfill$↓:\type{((t → t) → t) → t}$\label{def:lower}
\xe

The definition of continuized \ac{cfa} is a little more complicated. Informally,
(\ref{cfa}) says that \ac{cfa} takes a \textit{scopal} function and a
\textit{scopal} argument (in either order), unwraps the contained values and
performs \ac{fa} on them, while sequencing the quantificational parts from left-to-right.

\ex
\acf{cfa} (def.)\\
$m \ml{S} n ≔ λ k . m (λ x . n (λ y . k (x \ml{A} y)))$\\
\phantom{,}\hfill$\ml{S}:\begin{aligned}[t]
  &\type{(((a → b) → t) → t) → ((a → t) → t) → (b → t) → t}\\
  &\type{((a → t) → t) → (((a → b) → t) → t) → (b → t) → t}
  \end{aligned}$\label{cfa}
\xe

\todo[inline]{Show how this deals with quantifiers}

\subsection{Conjunction in continuation semantics}

Within continuation semantics, conjunction receives an extremely natural
definition, given in (\ref{def:conj}); \textit{and} takes two scope-takers as
$m$ and $n$, and a \textit{continuation argument} $k$, and feeds $k$ into $m$
and $n$. As shown in detail by \citeauthor{barker2002} this subsumes all of the
cases covered by \citeauthor{parteeRooth}'s generalized conjunction.

\ex
Conjunction in continuation semantics (def.)\\
$\eval{and} ≔ λ mnk . n k ∧ m k$\hfill$\type{((a → t) → t) → ((a → t) → t) → (a → t) → t}$
\xe

Note that, in the following, we'll abbreviate types of the form
$\type{(a → t) → t}$ as follows:

\ex
$\type{C a ≔ (a → t) → t}$
\xe



\todo[inline]{Give some examples}

\subsection{Accounting for the split-scope signature}

Continuation semantics accounts for the split-scope signature straightforwardly,
on the assumption that quantifiers can themselves be lifted (this is
independently necessary to account for inverse scope readings). We'll make a number of simplifying assumptions
here, but nothing in the analysis will crucially hinge upon them. For example,
we'll treat \ac{npi} \textit{any} simply as an existential quantifier, which is
licensed as long as it takes scope within the argument of \textit{refuse}.

The first step is to lift \textit{any European city} and \textit{any Asian
  city}, returning two higher-order scope-takers. Since \textit{and} can conjoin
anything scopal, it can conjoin the resulting meanings. The result is shown in
the figure below:

\begin{figure}[H]
  \centering
  \caption{Lift the \acp{npi} and conjoin}
\begin{forest}
  [{$λ l . l (λ k₂ . ∃y[ \ml{european-city} ∧ k₂ y])) ∧ l (λ k₁ . ∃x[\ml{asian-city} ∧ k₁ x])$\\\ac{fa}},draw=red
    [{$λ l . l (λ k₂ . ∃y[ \ml{european-city} ∧ k₂ y])$}
    [{$λ k₂ . ∃y[ \ml{european-city} ∧ k₂ y]$},edge label={node[midway,left,font=\scriptsize]{$\uparrow$}} [{any European city},roof]]]
    [{\acs{fa}}
      [{and}]
      [{$λ l . l (λ k₁ . ∃x[\ml{asian-city} ∧ k₁ x])$}
      [{$λ k₁ . ∃x[\ml{asian-city} ∧ k₁ x]$},edge label={node[midway,left,font=\scriptsize]{$\uparrow$}} [{any Asian city},roof]]]
    ]
  ]
\end{forest}
\end{figure}

The result of conjoining the lifted \acp{npi} is a higher-order scope taker. In
order to get this to compose with the rest of the embedded clause, we lift
\textit{visit} twice, and do higher-order \ac{cfa}. Informally, the result is
that the semantic contribution of \textit{visit} is distributed between the two
conjunctions. Next, we \textit{internally lower} the result, in order to set the
scope of the \acp{npi}.

\begin{figure}[H]
  \centering
  \caption{Compose the embedded clause and internally lower}
  \begin{forest}
    [{$λ l . \begin{aligned}[t]
        &l (∃y[ \ml{european-city} ∧ \ml{visit} y]))\\
        &∧ l (∃x[\ml{asian-city} ∧ \ml{visit} x])
        \end{aligned}$},draw=red
    [{...}
    [{\textsc{pro}$^{↑₂}$}]
    [{$λ l . \begin{aligned}[t]
        &l (λ k₂ . ∃y[ \ml{european-city} ∧ k₂ (\ml{visit} y)]))\\
        &∧ l (λ k₁ . ∃x[\ml{asian-city} ∧ k₁ (\ml{visit} x)])
        \end{aligned}$\\$⊛$}
      [{visit$^{\uparrow_{2}}$}]
      [{$λ l . \begin{aligned}[t]
          &l (λ k₂ . ∃y[ \ml{european-city} ∧ k₂ y]))\\
          &∧ l (λ
          k₁ . ∃x[\ml{asian-city} ∧ k₁ x])
          \end{aligned}$},fill=yellow]
    ]]]
  \end{forest}
\end{figure}

Now we compose the result with \texit{refuse} and the subject via \ac{sfa}, and lower the result.

\begin{figure}[H]
\centering
\caption{Compose the matrix clause and lower}
\begin{forest}
  [{$\ml{j refuse} ∃y[ \ml{european-city} ∧ \ml{j visit} y] ∧ ∃x[\ml{asian-city} ∧ \ml{visit} x]$},draw=red
  [{$⊛$}
    [{John$^{↑}$}]
    [{$⊛$}
      [{refuse$^{\uparrow}$}]
      [{$λ l . l (∃y[ \ml{european-city} ∧ \ml{visit} y])) ∧ l (∃x[\ml{asian-city} ∧ \ml{visit} x])$},fill=yellow]
    ]
  ]]
\end{forest}
\end{figure}

Zooming out, the graph of the semantic computation is given below:

\begin{figure}[H]
  \centering
  \begin{forest}
    [{$\type{t}$}
    [{$\type{C t}$},edge label={node[midway,left,font=\scriptsize]{$↓$}}
      [{John$^{↑}$}]
      [{$\type{C (e → t)}$}
        [{refused$^{↑}$}]
        [{$\type{C t}$}
        [{$\type{C (C t)}$},edge label={node[midway,left,font=\scriptsize]{$↓↓$}}
          [{\textsc{pro}}]
          [{$\type{C (C (e → t))}$\\$⊛$}
            [{$\type{C (C (e → e → t))}$\\visit$^{↑}$}]
            [{$\type{C (C e)}$}
              [{$\type{C (C e)}$} [{any European city},edge label={node[midway,left,font=\scriptsize]{$\uparrow$}}]]
              [{$\type{C (C e) → C (C e)}$}
                [{$\type{C a → C a → C a}$\\and}]
                [{$\type{C (C e)}$} [{any Asian city},edge label={node[midway,left,font=\scriptsize]{$\uparrow$}}]]
              ]
            ]
          ]
        ]
        ]
      ]
    ]]
  \end{forest}
\end{figure}

It's worth noting that \citet{hirschThesis} does briefly consider and ultimately
dismiss a similar solution, which involves type-lifting the \acp{npi}, and
allowing \acs{qr} of the resulting complex quantifier formed by
\textit{and}\sub{Q} to leave behind a higher type (i.e., a quantificational
trace), thus giving rise to semantic reconstruction. \citeauthor{hirschThesis}
ultimately rejects this solution, on the basis that such a derivation
disentangles syntactic position and scope, which has been argued to be
undesirable based on evidence that semantic reconstruction of a moved expression
feeds condition C. This consideration is completely irrelevant here, since
scope-taking does not logically entail \textit{movement} in continuation semantics.

\subsection{An unattested scope reading}

\ex
Some company hired a maid and a cook.\hfill\cmark $∃ > ∧$; \xmark $∧ > ∃$
\xe

\section{\textit{Only} and the split-scope signature}

\citet{hirschThesis} makes a parallel argument for split-scope for pre-DP
\textit{only}, intended to rule out the possibility that \textit{only} can
compose directly with a DP. The examples \citeauthor{hirschThesis} considers are
those like the following:

\ex
You're only required to read three books.\hfill $\ml{only} > ◻ > ∃$
\xe

In this section, I'll show that the analysis of
\textit{and} in continuation semantics (\citealt{barker2002}, \citealt[chapter 7]{barkerShan2015}), which involves re-executing a continuation, can be extended in a reasonably
straightforward fashion to \textit{only}, thus accounting for the
cross-categorial nature of \textit{only}, and \textit{predicting} split-scope by
virtue of the availability of \textit{lift}.

\subsection{Generalized \textit{only}}

% \ex
% $\ml{F} a ≔ (a,\set{a'|a' ∈ \ml{alt} a})$
% \xe

% \pex
% \a $a^{π} ≔ (a,\set{a})$
% \a
% $(x,\mathbb{x}) \ml{R} (y,\mathbb{y}) ≔ (x \ml{A} y,\set{x \ml{A} y|x ∈ \mathbb{x}, y ∈ \mathbb{y}})$
% \xe
%
Let's assume that \textit{three\sub{F} papers} denotes a pair, consisting of its
ordinary semantic value, and its focus-semantic value.

\ex
$\eval{three\sub{F} papers} = \left(\begin{aligned}[c]
    &λ k . ∃X[|X| = 3 ∧ \ml{papers} X],\\
    &\set{λ k . ∃X[|X| = n ∧ \ml{papers} X] | n ∈ \mathbb{N}}
    \end{aligned}\right)$\hfill$\type{\ml{F} (C a)}$
\xe

We can analyze \textit{only} as looking for a scope-taker associated with some
alternatives, and returning a scope-taker.

\ex
$\ml{only} (m,\mathbb{m}) ≔ λ k . m k ∧ ∀m' ∈ \mathbb{m}[(m k ↛ m' k) → ¬ (m' k)]$\\
\phantom{,}\hfill$\type{F (C a) → C a}$
\xe

% \begin{figure}[H]
%   \centering
% \caption{\texttt{fmap} lift through the focus applicative}
% \begin{forest}
%   [{$\left(\begin{aligned}[c]
%     &λl . l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k X]),\\
%     &\set{λ l . l (λ k . ∃X[|X| = n ∧ \ml{papers} X ∧ k X]) | n ∈ \mathbb{N}}
%     \end{aligned}\right)$},draw=red
%     [{...} [{three\sub{F} papers}]]
%   ]
% \end{forest}
% \end{figure}

% \begin{figure}[H]
%   \centering
%   \caption{Compose with \textit{only}}
%   \begin{forest}
%     [{$λ l . l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X]) ∧ ∀m' ∈ \ml{alts}[(l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k X]) ↛ m' l) → ¬ (m' l)]$},draw=red
%       [{only}]
%       [{$\left(\begin{aligned}[c]
%     &λl . l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k X]),\\
%     &\set{λ l . l (λ k . ∃X[|X| = n ∧ \ml{papers} X] ∧ k X) | n ∈ \mathbb{N}}
%     \end{aligned}\right)$},fill=yellow]
%     ]
%   \end{forest}
% \end{figure}

% \begin{figure}[H]
%   \centering
%   \caption{Compose embedded clause, and internally lower}
%   \begin{forest}
%     [{$λ l . l (∃X[|X| = 3 ∧ \ml{papers} X ∧ \ml{you read} X]) ∧ ∀m' ∈ \ml{alts}[(l (∃X[|X| = 3 ∧ \ml{papers} X ∧ \ml{you read} X]) ↛ m' l) → ¬ (m' l)]$}
%     [{$λ l . l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k (\ml{you read} X)]) ∧ ∀m' ∈ \ml{alts}[(l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X]) ↛ m' l) → ¬ (m' l)]$\\$\ml{S}_{2}$}
%       [{you$^{\uparrow_{2}}$}]
%       [{$\ml{S}_{2}$}
%         [{read$^{↑_{2}}$}]
%         [{$λ l . l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k X]) ∧ ∀m' ∈
%           \ml{alts}[(l (λ k . ∃X[|X| = 3 ∧ \ml{papers} X ∧ k X]) ↛ m' l) → ¬ (m' l)]$},fill=yellow]
%       ]
%     ]
%     ]
%   \end{forest}
% \end{figure}

\begin{figure}[H]
  \centering
  \caption{Graph of the derivation}
  \begin{forest}
    [{$\type{t}$}
    [{$\type{C t}$},edge label={node[midway,left,font=\scriptsize]{$\downarrow$}}
      [{$\type{C (C (t → t))}$\\require$^\uparrow$}]
      [{$\type{C t}$}
      [{$\type{C (C t)}$},edge label={node[midway,left,font=\scriptsize]{$\intLower$}}
        [{$\type{C (C e)}$\\you$^{↑₂}$}]
        [{$\type{C (C (e → t))}$}
          [{$\type{C (C (e → e → t))}$\\read$^{↑₂}$}]
          [{$\type{C (C e)}$}
            [{$\type{F (C a) → C a}$\\only}]
            [{$\type{F (C (C e))}$} [{$\type{F (C e)}$},edge label={node[midway,left,font=\scriptsize]{$\mathtt{fmap} \uparrow$}} [{three\sub{F} books},roof]]]
          ]
        ]
      ]
    ]]]
  \end{forest}
\end{figure}

\ex
$\begin{aligned}[t]
  &□ ∃X[|X| = 3 ∧ \ml{books} X ∧ \ml{you read} X]\\
  &∧ ∀p ∈ \ml{alts}[((□ ∃X[|X| = 3 ∧ \ml{books} X ∧ \ml{you read} X]) ↛ p) → ¬ p]\end{aligned}$
\xe

\printbibliography

\end{document}


%%% Local Variables:
%%% mode: latex
%%% TeX-engine: xetex
%%% TeX-master: t
%%% End:

% LocalWords:  Keny Niedzielski Schwarzchild Yasu Sudo
